---
title: "Covid19TARShiny"
author: ""
output: 
        flexdashboard::flex_dashboard:
        orientation: columns, row
        vertical_layout: fill
---

```{r setup, include=FALSE}
library(ggplot2)
library(plotly)
library(plyr)
library(flexdashboard)
library(tidyverse)
library(ROCR)
library(MASS) # for Pima data sets
library(plotROC)

view()

# create some data
dataTARCovid19Adiniko=read.csv("D:/fileniko/TA_ADINIKO_0641716/aplikasi/COVID19_line_list_data")
data<- dataTARCovid19Adiniko

head(dataTARCovid19Adiniko)
View(dataTARCovid19Adiniko)
colnames(dataTARCovid19Adiniko)
dim(dataTARCovid19Adiniko)
str(dataTARCovid19Adiniko)

str(dataTARCovid19Adiniko)
#$ gender        : num  1 0 1 0 1 0 1 1 1 1 ...

table(data$death)
table(data$recovered)
table(dataTARCovid19Adiniko$death)
dataTARCovid19Adiniko$death_dummy=as.integer(dataTARCovid19Adiniko$death != 0)
table(dataTARCovid19Adiniko$death_dummy)
dataTARCovid19Adiniko$recovered_dummy=as.integer(dataTARCovid19Adiniko$recovered !=0)
table(dataTARCovid19Adiniko$recovered_dummy)
table(dataTARCovid19Adiniko$visiting.Wuhan, dataTARCovid19Adiniko$recovered_dummy)
table(dataTARCovid19Adiniko$from.Wuhan, dataTARCovid19Adiniko$recovered_dummy)
view(data)

nrow(dataTARCovid19Adiniko[is.na(dataTARCovid19Adiniko$gender) | is.na(dataTARCovid19Adiniko$age),])
dataTARCovid19Adiniko[is.na(dataTARCovid19Adiniko$age) | is.na(dataTARCovid19Adiniko$gender),]
nrow(dataTARCovid19Adiniko)
data <- dataTARCovid19Adiniko[!(is.na(dataTARCovid19Adiniko$gender) | is.na(dataTARCovid19Adiniko$age)),]
nrow(data)

#gender
data[data$gender == "female",]$gender <- 0
data[data$gender == "male",]$gender <- 1
data$gender <- as.numeric(data$gender)

#1. satistika deskriptif

attach(data)
table(death_dummy)
table(gender)
table(recovered_dummy)
table(visiting.Wuhan)
table(from.Wuhan)

summary(data)
xtabs(~ death_dummy + gender,data=data)
xtabs(~ death_dummy + age,data=data)
xtabs(~ death_dummy + recovered_dummy,data=data)
xtabs(~ death_dummy + visiting.Wuhan,data=data)
xtabs(~ death_dummy + from.Wuhan,data=data)

#2. Tabulasi Silang
ct <- table(gender,death_dummy)
ftable(ct)
ct1 <- table(age,death_dummy)
ftable(ct1)
ct2 <- table(recovered_dummy,death_dummy)
ftable(ct2)
ct3 <- table(visiting.Wuhan,death_dummy)
ftable(ct3)
ct4 <- table(from.Wuhan,death_dummy)
ftable(ct4)


#3. regresi logistik gender, usia, recovered dan kematian
library(oddsratio)
library(ggplot2)
library(cowplot)
library(base)
library(dplyr)
library(foreign)
library(nnet)
library(ordinal)
library(MASS)
library(pscl)
library(lmtest)
library(zoo)
library(haven)

#3. Uji Proposisi Gender terhadap kematian
#gender vs death
str(data)
male=subset(data, data$gender==1)
female=subset(data, data$gender==0)

dim_male = dim(male)
dim_male

dim_female = dim(female)
dim_female

sum_male = table(male$death_dummy)
sum_male

sum_female = table(female$death_dummy)
sum_female

#recovered vs death
notsurvive=subset(data, data$recovered_dummy==0)
survive=subset(data, data$recovered_dummy==1)

dim_notsurvive = dim(notsurvive)
dim_notsurvive

dim_survive = dim(survive)
dim_survive

sum_notsurvive = table(notsurvive$death_dummy)
sum_notsurvive

sum_survive = table(survive$death_dummy)
sum_survive

#visit wuhan vs death
notvisiting=subset(data, data$visiting.Wuhan=="0")
visiting=subset(data, data$visiting.Wuhan=="1")

dim_notvisiting = dim(notvisiting)
dim_notvisiting

dim_visiting = dim(visiting)
dim_visiting

sum_notvisiting = table(notvisiting$death_dummy)
sum_notvisiting

sum_visiting = table(visiting$death_dummy)
sum_visiting

#from wuhan vs death
notfromwuhan=subset(data, data$from.Wuhan=="0")
fromwuhan=subset(data, data$from.Wuhan=="1")

dim_notfromwuhan = dim(notfromwuhan)
dim_notfromwuhan

dim_fromwuhan = dim(fromwuhan)
dim_fromwuhan

sum_notfromwuhan = table(notfromwuhan$death_dummy)
sum_notfromwuhan

sum_fromwuhan = table(fromwuhan$death_dummy)
sum_fromwuhan

# uji perbedaan proporsi antar 2 grup dengan menggunakan z-test
# Ho : p death_rate_male is egual to p death_rate_female
# Ha : p death_rate_male is not equal to p death_rate_female

#prop gender to death
prop.test(x = c(44, 14), n = c(476, 349),alternative = "two.sided", conf.level = .95, correct = FALSE)

#prop recovered to death
prop.test(x = c(58, 0), n = c(682, 143),alternative = "two.sided", conf.level = .95, correct = FALSE)

#prop visiting.Wuhan to death
prop.test(x = c(57, 1), n = c(677, 148),alternative = "two.sided", conf.level = .95, correct = FALSE)

#prop from.Wuhan to death
prop.test(x = c(24, 34), n = c(671, 150),alternative = "two.sided", conf.level = .95, correct = FALSE)

#Uji T

dead=subset(data, data$death_dummy ==1)
alive=subset(data, data$death_dummy == 0)
# Ho : miu age_dead is equal to miu age_alive
# H1 : miu age_dead  is not equal to age_alive

t.test(dead$age, alive$age, alternative = "two.sided")



#MOdel REGRESI LOGIT
#Model1
logistik1 <- glm(data$death_dummy ~ data$gender +data$age + data$recovered_dummy +data$visiting.Wuhan + data$from.Wuhan, data = data, family = binomial(link="logit"))

#Uji Overall/simultan
pR2(logistik1)
qchisq(0.95,7)

#Uji Parsial
summary(logistik1)

#Model2
logistik2=glm(data$death_dummy ~ data$gender +data$age +data$recovered_dummy + data$from.Wuhan, family=binomial (link="logit"), data = data)
summary(logistik2)

#Model3
logistik3=glm(data$death_dummy ~ data$gender +data$age + data$visiting.Wuhan  +data$from.Wuhan, family=binomial (link="logit"),data = data)
summary(logistik3)

#Model4
logistik4=glm(data$death_dummy ~ data$gender +data$age + data$from.Wuhan, family=binomial (link="logit"), data = data)
summary(logistik4)

Model=c("Model 1","Model 2","Model 3","Model 4")
AIC = c(logistik1$aic,logistik2$aic,logistik3$aic, logistik4$aic)

Tabel_Kriteria=data.frame(Model,AIC)
Tabel_Kriteria

#Koefisien Determinan
pR2(logistik2)
qchisq(0.95,6)

library(ResourceSelection)
hl <- hoslem.test(logistik2$y, fitted(logistik2))
hl

library(aod)
waldtest(logistik2, logistik1, test = "Chisq")


```

# gender\~death

## Column {data-width="250"}

### geom_jitter gender\~death

```{r}
a <- ggplot(data, aes(gender, death_dummy))
a + geom_jitter(aes(colour = factor(gender), shape = factor(death_dummy), size = gender))

```

## Row {data-width="250"}

### geom pointsmooth gender~death

```{r}
ggplot(data, aes(gender, as.numeric(death_dummy)-1 )) +
  geom_point() +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = "binomial"))

```

## Row {data-width="200"}

### hist gender
```{r}
hist(data$gender)
```


# age~death

## Column {data-width="650"}

### geom_jitter age~death

```{r}
a <- ggplot(data, aes(age, death_dummy))
a + geom_jitter(aes(colour = factor(age), shape = factor(death_dummy), size = age))

```

## Column {data-width="350"}

### geom pointsmooth age~death

```{r}
ggplot(data, aes(age, as.numeric(death_dummy)-1 )) +
  geom_point() +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = "binomial"))

```

### hist age

```{r}
hist(data$age)
```

# recover~death

## Column {data-width="650"}

### geom_jitter recovered~death

```{r}
a <- ggplot(data, aes(recovered_dummy, death_dummy))
a + geom_jitter(aes(colour = factor(recovered_dummy), shape = factor(death_dummy), size = recovered_dummy))

```

## Column {data-width="350"}

### geom pointsmooth recovered~death

```{r}
ggplot(data, aes(recovered_dummy, as.numeric(death_dummy)-1 )) +
  geom_point() +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = "binomial"))

```

### hist recovered

```{r}
hist(data$recovered_dummy)
```

# visitWuhan~death

## Column {data-width="650"}

### geom_jitter visitingWuhan~death

```{r}
a <- ggplot(data, aes(visiting.Wuhan, death_dummy))
a + geom_jitter(aes(colour = factor(visiting.Wuhan), shape = factor(death_dummy), size = visiting.Wuhan))

```

## Column {data-width="350"}

### geom pointsmooth visiringWuhan~death

```{r}
ggplot(data, aes(visiting.Wuhan, as.numeric(death_dummy)-1 )) +
  geom_point() +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = "binomial"))

```

### hist visitingWuhan

```{r}
hist(data$visiting.Wuhan)
```

# fromWuhan~death

## Column {data-width="650"}

### geom_jitter fromWuhan\~death

```{r}
a <- ggplot(data, aes(from.Wuhan, death_dummy))
a + geom_jitter(aes(colour = factor(from.Wuhan), shape = factor(death_dummy), size = from.Wuhan))

```

## Column {data-width="350"}

### geom pointsmooth fromWuhan~death

```{r}
ggplot(data, aes(from.Wuhan, as.numeric(death_dummy)-1 )) +
  geom_point() +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = "binomial"))

```

### hist fromWuhan

```{r}
hist(data$from.Wuhan)
```

```{r}
library(caret)
library(InformationValue)
library(ISLR)

#Splitdata
#make this example reproducible
set.seed(1)

#Use 90% of dataset as training set and remaining 10% as testing set
sample <- sample(c(TRUE, FALSE), nrow(data), replace=TRUE, prob=c(0.90,0.10))
train <- data[sample, ]
test <- data[!sample, ]  

#fit logistic regression model
model <- glm(death_dummy ~ gender + age + recovered_dummy + from.Wuhan, family="binomial", data=train)
model

#disable scientific notation for model summary
options(scipen=999)

#view model summary
summary(model)
summary(data)

pscl::pR2(model)["McFadden"]

caret::varImp(model)

#calculate VIF values for each predictor variable in our model
car::vif(model)

#define two individuals
new <- data.frame(gender = c(0, 1), age = 48,recovered_dummy = c(0, 1),from.Wuhan = c(0, 1))

#predict probability of defaulting
predict(model, new, type="response")

#calculate probability of default for each individual in test dataset
predicted <- predict(model, test, type="response")
predicted

library(InformationValue)

#convert defaults from "Yes" and "No" to 1's and 0's
test$death_dummy <- ifelse(test$death_dummy=="Death", 1, 0)

#find optimal cutoff probability to use to maximize accuracy
optimal <- optimalCutoff(test$death_dummy, predicted)[1]
optimal

table(test$death_dummy)

library(caret)
#confusionMatrix(test$Y, test$predicted)

#Confusion Matrix
table(test$death_dummy, predicted > 0.10)
table(test$death, predicted > 0.10)
#confusionMatrix(test$death_dummy, predicted)

#calculate sensitivity
sensitivity(test$death_dummy, test$predicted)
sensitivity <- (75/(75+0))
sensitivity

#calculate specificity
specificity(test$death_dummy, predicted)
specificity <- (6/(6+0))
specificity

#calculate specificity
precision(test$death_dummy, predicted)
precision <- (75/(75+9))
precision

#calculate total misclassification error rate
misClassError(test$death_dummy, predicted, threshold=optimal)

# Accuracy of model:
(75+6)/(75+9+0+6)

# combine linear predictor and known truth for training and test datasets into one data frame
df <- rbind(data.frame(predictor = predict(model, data),known.truth = data$death_dummy,model = "train"),
data.frame(predictor = predict(model, data),known.truth = data$death_dummy, model = "test"))

#odds rasio
exp(coef(model))
#manual 
#GENDER ODDS
odds=glm(data$death_dummy ~ data$gender, family=binomial)
or_glm(data=data, model=odds)
odds

# pi(X=1)peluang mengalami kematian akibat covid dari pasien pria
m=2.71828^(-3.1751+0.8908*1)/(1+2.71828^(-3.1751+0.8908*1))
m
# [1] 0.08461769
# pi(X=0)peluang mengalami kematian akibat covid dari pasien wanita
f=2.71828^(-3.1751+0.8908*0)/(1+2.71828^(-3.1751+0.8908*0))
f

odds_male=m/(1-m)
odds_female=f/(1-f)
odds_ratio1=odds_male/odds_female
odds_ratio1

#AGE ODDS
library(oddsratio)
odds=glm(data$death_dummy ~ data$age, family=binomial)
or_glm(data=data, model=odds)
odds

# peluang mengalami kematian akibat covid dari pasien usia 30
usia30 = 2.71828^(-7.31275+0.07981 * 30)/(1+2.71828^(-7.31275+0.07981 *30))
usia30
# peluang mengalami kematian akibat covid dari pasien usia 70
usia31 = 2.71828^(-7.31275+0.07981 * 31)/(1+2.71828^(-7.31275+0.07981 *31))
usia31

odds_usia30=usia30/(1-usia30)
odds_usia31=usia31/(1-usia31)
odds_ratio2=odds_usia30/odds_usia31
odds_ratio2
```
# Prdic,ROCR

## Rows {data-width="350"}

### Prediksi

```{r}
test$data<-predict(model, type = "response", newdata = test)
ggplot(test, aes(x=data)) +
  geom_density(lwd=0.5)+
  labs(title = "Distribusi Prob Data Prediksi Kematian Pasien") +
  theme_minimal()
```
# ROCR

## Rows {data-width="350"}

### ROCR

```{r}
# the aesthetic names are not the most intuitive
# `d` (disease) holds the known truth
# `m` (marker) holds the predictor values 
ggplot(df, aes(d = known.truth, m = predictor, color = model)) + 
  geom_roc()
```


## Rows {data-width="350"}

### AUROCR
```{r}
#plotroc
plotROC(test$death, predicted)

```
